A simple and efficient algorithm for computing approximate Nash equilibria in sequential games with imcomplete information

### 2015-07-28 paper.tex
fixups before final arXiv upload (version 1)

\maketitle
\begin{abstract}We present a simple primal-dual algorithm for
  computing approximate Nash equilibria in two-person zero-sum
  sequential games with  incomplete information and perfect recall
  (like Texas Hold'em poker). Our algorithm only performs basic
  iterations (i.e iterations involving no calls to external
  first-order oracles, etc.)
  and is applicable to a broad class of two-person zero-sum games
  including simultaneous games and sequential games with incomplete
  information and perfect recall. The applicability to the latter kind
  of games is thanks to the sequence-form representation
  \cite{koller1992complexity} which allows us to encode any such game
  as a matrix game with polyhedral strategy profiles. We prove that
  the number of iterations needed to produce a Nash equilibrium with a
  given precision is inversely proportional to the
  precision. We also present experimental results on simulated and
  real games (Kuhn poker).
\end{abstract}

@@

and the dual LCP
\begin{equation}
  \begin{aligned}
    \text{maximize}-\langle e_2, q\rangle,\hspace{.5em}\text{subject
      to: }(x,q) \in \mathbb{R}^{n_1} \times \mathbb{R}^{l_2},  x \ge
    0, E_1x = e_1, A^Tx + E_2^Tq \geq 0.
  \end{aligned}
  \label{eq:dual_pb}
\end{equation}
The vectors $p = (p_0, p_1, ..., p_{l_2 - 1}) \in \mathbb{R}^{l_2}$
and $q = (q_0, q_1, ..., q_{l_1 - 1}) \in \mathbb{R}^{l_1}$ are dual
variables. 
$A$ is the \textit{payoff matrix} and each $E_k$ is a matrix whose
entries are $-1$, $0$ or $1$, with exactly 1 entry equal -1 per row, and each $e_k$ is a vector of the form
$(1, 0, ..., 0)$.

\begin{figure}[!h]
\begin{minipage}{.65\linewidth}
%\caption*{\textbf{Sparse-Variation.}
Note that the LCPs above have the equivalent saddle point formulation
\begin{equation}
  \underset{y \in Q_2}{\text{minimize}}\text{ }\underset{x \in
    Q_1}{\text{maximize}}\text{ }\langle x, Ay\rangle,
  \label{eq:gilpin}
\end{equation}
where the \textit{convex polytope} (by which we mean a bounded convex
polyhedron)
\begin{equation}
  Q_k := \{z \in \mathbb{R}^{n_k}_+ |\text{ }E_kz = e_k\} \subseteq
  [0, 1]^{n_k}
\label{eq:polytope}
\end{equation}
is the the strategy profile of player $k$ in this representation.
At a feasible point $(y, p, x, q)$ for the LCPs, the \textit{primal-dual gap}
$\tilde{G}(y, p, x, q)$ of these primal-dual pair is given
by\footnote{The inequality being due to \textit{weak duality}.}
\begin{eqnarray}
  \begin{split}
  0 &\le \tilde{G}(y, p, x, q) := \langle e_1, p\rangle - (-\langle
  e_2, q\rangle) = \langle e_1, p\rangle + \langle
  e_2, q\rangle\\
  &= G(x, y) := \mathrm{max}\{\langle u, Ay\rangle - \langle x, Av\rangle |
(u,v) \in Q_1 \times Q_2\}.
\end{split}
  \label{eq:dgap}
\end{eqnarray}
%}
\end{minipage}
\hspace{.5em}
\begin{minipage}{.4\linewidth}
  \includegraphics[width=1\linewidth]{grain4.png}
  \caption{A convex polytope}
  \label{fig:roi}
\end{minipage}
\end{figure}

@@

\paragraph{A note about matrix games on simplexes.} It should be
noted
that any matrix $A \in \mathbb{R}^{n_1 \times n_2}$ specifies a matrix
  game with payoff matrix $A$, for which each player's strategy
profile is a simplex; this simplex can be written in the form
\eqref{eq:polytope} by taking $E_k := (1, 1, ..., 1) \in
\mathbb{R}^{1 \times n_k}$ and $e_k = 1 \in \mathbb{R}^1$. Thus every
matrix game on simplexes can be seen as a sequential game.  Thus the
results presented in this manuscript can be trivially applied such
games in particular. Here, the polyhedral strategy profiles $Q_k$
defined in \eqref{eq:polytope} reduce to simplexes $\Delta_{n_k}$,
and the primal-dual gap function $G(x,y)$ writes
\begin{eqnarray}G(x, y) =
\mathrm{max}\{\langle u, Ay\rangle - \langle x, Av\rangle | (u,v) \in
\Delta_{n_1} \times \Delta_{n_2}\} = \underset{0 \le i <
  n_1}{\text{max }}(Ay)_i - \underset{0 \le j < n_2}{\text{min
}}(A^Tx)_j.
\label{eq:mg_pd}
\end{eqnarray}

@@

\subsection{Our contribution}
One cannot directly attack the LCPs \eqref{eq:primal_pb} and
\eqref{eq:dual_pb} via a traditional primal-dual algorithm
(for example \cite{chambolle2010,chambolle2014ergodic}) because
computing the orthogonal projections $\proj_{Q_k}$ is very difficult
(in fact, such subproblems would have to be solved
iteratively\footnote{An ``exception to the rule'' is the case where
  the $Q_k's$ are simplexes, so that the projections can be computed
  exactly using \cite{duchi2008efficient}.}). Also,
the primal-dual gap might explode even at points arbitrarily close to
the set of feasible points, leaving the algorithm with no indication
whatsoever, on whether progress is being made or not. So need to way to
\begin{itemize}
\item[--] avoid having to compute the projections $\proj_{Q_k}$,
\item[--] have control over how far we are from the set of equilibria and
  avoid infinite (and thus non-informative) primal-dual gaps.
\end{itemize}

Developing on an alternative notion of approximate equilibria (see
Definition \ref{thm:cool_notion})
homologous to that presented in Definition \ref{thm:approx_nash}, we
device a primal-dual algorithm (Algorithm \ref{Tab:algo}) for
computing approximate Nash equilibria in sequential two-person
zero-sum games with incomplete information and perfect recall. We also
prove (Theorem \ref{thm:pd}) that --in an ergodic sense-- the number of
iterations required by the algorithm to produce an approximation
equilibrium to a precision $\epsilon$ is $\mathcal{O}(1/\epsilon)$,
with explicit values for the constants involved this worst-case cost.
These contributions will be elaborated in section \ref{sec:gsp}.
%%  The source of approximation in our scheme
%% is that we explicitly control how close we are to the feasible set
%% $Q_1 \times Q_2$.

\section{Related work}
\label{sec:related_work}
We present a selection of algorithms that is representative of the
efforts that have been made in the literature to compute Nash
$\epsilon$-equilibria for two-person zero-sum games with incomplete
information like Texas Hold'em poker, etc.
First and foremost, let us note that for the class of games considered
here (sequential games with incomplete information), the LCPs
\eqref{eq:primal_pb} and \eqref{eq:dual_pb} are exceedingly larger
than what state-of-the-art LCP and interior-point solvers can
handle. See for example \cite{hoda2010smoothing,gilpinfirst}.

In \cite{hoda2010smoothing}, a nested iterative procedure using the
Excessive Gap Technique (EGT) \cite{nesterov2005excessive} was used to
to solve the equilibrium problem \eqref{eq:gilpin}.
The authors reported a $\mathcal{O}(1/\epsilon)$ convergence rate
(which derives from the general EGT theory) for the outer-most
iteration loop.
\cite{gilpinfirst} proposed a modified verson of the technqiues in
\cite{hoda2010smoothing} and  proved a $\mathcal{O}\left(\left(\|A\| /
\delta\right) ln\left(1 / \epsilon\right)\right)$ convergence rate in
terms of the number of calls made to a first-order oracle. Here
$\delta = \delta(A, E_1, E_2, e_1, e_2) > 0$ is a certain
\textit{condition number} for the game. The crux of their technique was to
observe that \eqref{eq:gilpin} can further be written a the minimization of
the primal-dual gap function $G(x, y)$ (defined in \eqref{eq:dgap})
for the game\footnote{The minimizers of $G$ are precisely the
  equilibria of the game.}, viz
\begin{eqnarray}
\mathrm{minimize}\{G(x,y)|(x,y) \in Q_1 \times Q_2\},
\end{eqnarray}
and then show there exists a scalar
$\delta > 0$ such that for any pair of realization plans $(x, y) \in Q_1 \times Q_2$,
\begin{eqnarray}
\text{``distance between }(x, y)\text{ and the set of
equilibria'' } \le G(x, y)/\delta.
\end{eqnarray}
Their
algorithm is then derived by iteratively applying Nesterov smoothing \cite{nesterov2005a}
with a geometrically decreasing sequence of tolerance levels
$\epsilon_{n+1} = \epsilon_n / \gamma$ (with $\gamma > 1$)  $G$. It
should be noted however that
\begin{itemize}
\item[--] The constant $\delta > 0$ can be arbitrarily small, and so
  the factor $\|A\| / \delta$ in the $\mathcal{O}\left(\left(\|A\| /
\delta\right) ln\left(1 / \epsilon\right)\right)$ convergence rate can
be arbitrarily large for ill-conditioned games.
\item[--] The reported linear convergence rate is not in terms of
  basic operations (addition, multiplication, matvec, clipping, etc.),
  but in terms of the number of calls to a first-order oracle. Most
  notably, the complicated projections $\proj_{Q_k}$ are applied at
  each iteration.%%  Of
  %% course, the existence of a linear algorithm for solving
  %% \eqref{eq:gilpin} is improbable, as the usual strong-convexity
  %% conditions are absent.

@@

The primal-dual algorithm first developed in \cite{chambolle2010}, was
proposed can be used to solve matrix games on
simplexes as was done in \cite{chambolle2014ergodic}, for example.
It should be stressed that such matrix games are
considerably simpler than the games considered here. Indeed, the
authors in \cite{chambolle2014ergodic} use the fact that computing the
orthogonal projection of a point onto a simplex can be done in linear
time as in \cite{duchi2008efficient}. In contrast, no such efficient
algorithm is known nor is likely to exist for the polyhedra $Q_k$
defined in \eqref{eq:polytope}, the strategy profiles for players in
the games considered here. It should however be noted that such this
projection can still be done iteratively using, for example, the
algorithm in proposition 4.2 of \cite{combettes2010dualization} or the
algorithms developed in \cite{tran2015splitting}. Unfortunately, as
with any nested iterative scheme, one would have to solve this
sub-problem with finer and finer precision.

@@

\begin{proof}
Note that $dom(g_1) = S_1$ and $dom(g_2) = S_2$ and observe that since the
equilibria of the LCPs \eqref{eq:primal_pb} and \eqref{eq:dual_pb}, and
the equilibria of GSP($\Psi_1$, $\Psi_2$, $g_1$, $g_2$) are contained
in $S_1 \times S_2$, it suffices to show that at any point $(y, p, x,
q) \in S_1 \times S_2$, the primal-dual gap between the primal
LCP \eqref{eq:primal_pb} and the dual LCP \eqref{eq:dual_pb} equals
the primal-dual gap of GSP($\Psi_1$, $\Psi_2$, $g_1$, $g_2$).
Indeed, the unconstrained objective in \eqref{eq:primal_pb}, say
$a(x,y)$, can be computed as
\begin{eqnarray*}
  \begin{aligned}
    a(y,p) &= \langle e_1,p\rangle + i_{y\ge 0} + i_{-Ay + E_1^Tp \ge 0} +
    i_{E_2y = e_2}\\
    &= g_1(y,p) + \underset{x' \geq
      0}{\text{max}}\text{ }\langle x',Ay - E_1^Tp\rangle +
    \underset{q'}{\text{max}}\text{ }\langle q',E_2y - e_2\rangle\\
    &= g_1(y,p) + \underset{x',
      q'}{\text{max}}\text{ }\langle x',Ay\rangle - \langle x',
    E_1^Tp\rangle + \langle q',E_2y\rangle -
    (i_{x' \ge 0} + \langle e_2,q\rangle)\\
    &= g_1(y,p)
      - \underset{x',q'}{\text{min}}\text{ }\Psi_2(y, p, x', q') + g_2(x',
      q') = g_1(y,p)
      - \underbrace{\underset{x',q'}{\text{min}}\text{
        }\hat{\Psi}_2(y, p, x', q')}_{\phi_2(y,p)}\\
      &= g_1(y, p) - \phi_2(y, p).
  \end{aligned}
  \label{eq:a}
\end{eqnarray*}

@@

The interested reader should refer to \cite{combettes2011proximal} and
the references therewithing, for a more elaborate exposition on
proximal calculus and its use in modern convex-optimization.

\subsection{A more comfortable notion of approximate Nash equilibrium}
Given a scalar $\epsilon > 0$ and a function $f:\mathbb{R}^n
\rightarrow [-\infty,+\infty]$, the $\epsilon$-enlarged subgradient (or
$\epsilon$-subgradient, for short) of $f$ is the set-valued function
\begin{eqnarray}
\partial_\epsilon f: \mathbb{R}^n \rightarrow
2^{\mathbb{R}^n}, \hspace{.5em} x \mapsto \{s \in \mathbb{R}^n | f(z)
\ge f(x) + \langle s, z - x\rangle - \epsilon,\forall z \in
\mathbb{R}^n\}.
\end{eqnarray}

The idea behind $\epsilon$-subgradients is the following. Say we wish
to minimize the function $f$. Replace the usual necessary
condition ``$0 \in \partial f(x)$'' for the optimality of $x$ with the
weaker condition ``$\partial_\epsilon f(x)$ contains a sufficiently
small vector $v$''. In fact, it is easy to see that, for each
point $x \in \mathbb{R}^n$, we have $\underset{\epsilon \rightarrow
  0^+}{\text{lim }}\partial_\epsilon f(x) = \partial f(x)$.

@@

\subsection{The proposed algorithm}
\label{sec:algo}
We now derive and establish the theoretical properties of the
algorithm which is the main object of this manuscript.

\begin{algorithm}
\caption{Primal-dual algorithm for computing approximate Nash
  Equilbria in two-person zero-sum games with incomplete information
  and perfect recall}
\label{Tab:algo}
\begin{algorithmic}[1]
\Require $\epsilon > 0$; $(y^{(0)},p^{(0)},x^{(0)},q^{(0)})
\in \mathbb{R}^{n_2}
  \times \mathbb{R}^{l_1} \times \mathbb{R}^{n_1} \times
  \mathbb{R}^{l_2}$.
\Ensure A Nash $(\epsilon,0)$-equilibrium
$({y^*},{p^*},{x^*},{q^*}) \in S_1 \times S_2$ for
the GSP \eqref{eq:unconstrained_pb}.
\State  $\lambda \leftarrow 1/\|K\|$, ${v}_a^{(0)} \leftarrow 0$, $k
\leftarrow 0$
\While{$\|{v}_{a}^{(k)}\|/k \ge \epsilon$}
\State $y^{(k + 1)} \leftarrow (y^{(k)} - \lambda (A^Tx^{(k)} +
E_2^Tq^{(k)}))_+$, \hspace{.5em}$p^{(k+1)} \leftarrow p^{(k)} -
\lambda(e_1-E_1x^{(k)})$
\State $x^{(k + 1)} \leftarrow (x^{(k)} + \lambda (Ay^{(k+1)} -
E_1^Tp^{(k+1)}))_+$, \hspace{.5em}$\Delta x^{(k+1)} \leftarrow
x^{(k+1)}-x^{(k)}$
\State $\Delta q^{(k+1)} \leftarrow \lambda (E_2y -
e_2)$, \hspace{.5em}$q^{(k+1)} \leftarrow q^{(k)} + \Delta q^{(k+1)}$
\State $y^{(k+1)} \leftarrow y^{(k+1)} - \lambda (A^T\Delta x^{(k+1)}
+ E_2^T\Delta q^{(k+1)})$, \hspace{.5em}$\Delta y^{(k+1)} \leftarrow
y^{(k+1)}-y^{(k)}$
\State $p^{(k+1)} \leftarrow p^{(k+1)} + \lambda E_1\Delta x^{(k+1)}$,
\hspace{.5em} $\Delta p^{(k+1)} \leftarrow p^{(k+1)}-p^{(k)}$
\State ${v}_a^{(k+1)} \leftarrow {v}_a^{(k)} +
\frac{1}{\lambda}(\Delta
y^{(k+1)},\Delta p^{(k+1)},\Delta x^{(k+1)},\Delta q^{(k+1)})$
\State $k \leftarrow k + 1$
\EndWhile
\end{algorithmic}
\end{algorithm}

\begin{theorem}[Ergodic $\mathcal{O}(1/\epsilon)$ convergence]
Let $d_0$ be the orthogonal distance between the starting point
$(y^{(0)},p^{(0)},x^{(0)},q^{(0)})$ of Algorithm \ref{Tab:algo} and the
set of equilibria for the GSP \eqref{eq:unconstrained_pb}.
Then given any $\epsilon > 0$, there exists an index
$k_0 \le \frac{2d_0\|K\|}{\epsilon}$ such that after $k_0$ iterations
the algorithm produces the following a quadruplet
$(y^{k_0},p^{k_0},x^{k_0},q^{k_0})$ and a vector $v^{k_0}$ such that
\begin{eqnarray}
\|v^{k_0}\| \le \epsilon \text{ and}\hspace{.5em} v^{k_0} \in
\partial[\hat{\Psi}_1(., ., x^{k_0}, q^{k_0}) +
  \hat{\Psi}_2(y^{k_0}, p^{k_0}, ., .)](y^{k_0},p^{k_0},x^{k_0},q^{k_0}).
\end{eqnarray}

@@

\section{Concluding remarks}
Making use of the sequence-form representation
\cite{koller1992complexity,von1996efficient,vonequilibrium}, we have
deviced a simple and efficient primal-dual algorithm for computing
Nash equilibria in two-person zero-sum sequential games with
incomplete information (like Texas Hold'em, etc.). Our algorithm is
simple to implement, with a low constant cost per iteration, and
enjoys a rigorous convergence theory with a proven
$\mathcal{O}(1/\epsilon)$ convergence in terms of basic operations
(matvec products, clipping, etc.), to a Nash
$(\epsilon,0)$-equilibrium of the game. Equilibrium problems are
saddle-point convex-concave problems, and as
such a natural choice for algorithms for solving them would be in the
family of primal-dual algorithms. We believe such primal-dual
schemes will receive more attention in the algorithmic game theory
community in future.

>>>>

\maketitle
\begin{abstract}We present a simple primal-dual algorithm for
  computing approximate Nash equilibria in two-person zero-sum
  sequential games with  incomplete information and perfect recall
  (like Texas Hold'em poker). Our algorithm only performs basic
  iterations (i.e matvec multiplications, clipping, etc., and no calls
  to external first-order oracles, no matrix inversions, etc.)
  and is applicable to a broad class of two-person zero-sum games
  including simultaneous games and sequential games with incomplete
  information and perfect recall. The applicability to the latter kind
  of games is thanks to the sequence-form representation
  \cite{koller1992complexity} which allows us to encode any such game
  as a matrix game with convex polytopial profiles. We prove that
  the number of iterations needed to produce a Nash equilibrium with a
  given precision is inversely proportional to the
  precision. We also present experimental results on simulated and
  real games (Kuhn poker).
\end{abstract}

@@

and the dual LCP
\begin{equation}
  \begin{aligned}
    \text{maximize}-\langle e_2, q\rangle,\hspace{.5em}\text{subject
      to: }(x,q) \in \mathbb{R}^{n_1} \times \mathbb{R}^{l_2},  x \ge
    0, E_1x = e_1, A^Tx + E_2^Tq \geq 0.
  \end{aligned}
  \label{eq:dual_pb}
\end{equation}
The vectors $p = (p_0, p_1, ..., p_{l_2 - 1}) \in \mathbb{R}^{l_2}$
and $q = (q_0, q_1, ..., q_{l_1 - 1}) \in \mathbb{R}^{l_1}$ are dual
variables. 
$A$ is the \textit{payoff matrix} and each $E_k$ is a matrix whose
entries are $-1$, $0$ or $1$, with exactly 1 entry per row which
equals $-1$ except for the first whose whose first entry is $1$ and all
the others are $0$. Each of the vectors $e_k$ is of the form $(1, 0, ..., 0)$.

\begin{figure}[!h]
\begin{minipage}{.65\linewidth}
%\caption*{\textbf{Sparse-Variation.}
Note that the LCPs above have the equivalent saddle point formulation
\begin{equation}
  \underset{y \in Q_2}{\text{minimize}}\text{ }\underset{x \in
    Q_1}{\text{maximize}}\text{ }\langle x, Ay\rangle,
  \label{eq:gilpin}
\end{equation}
where the \textit{convex polytope} (by which we mean a bounded convex
polyhedron)
\begin{equation}
  Q_k := \{z \in \mathbb{R}^{n_k}_+ |\text{ }E_kz = e_k\} \subseteq
  [0, 1]^{n_k}
\label{eq:polytope}
\end{equation}
is identified with the strategy profile of player $k$ in the
sequence-form representation. At a feasible point $(y, p, x, q)$ for
the LCPs, the \textit{primal-dual gap} $\tilde{G}(y, p, x, q)$ of
these primal-dual pair is given by\footnote{The inequality being due
  to \textit{weak duality}.}
\begin{eqnarray}
  \begin{split}
  0 &\le \tilde{G}(y, p, x, q) := \langle e_1, p\rangle - (-\langle
  e_2, q\rangle) = \langle e_1, p\rangle + \langle
  e_2, q\rangle\\
  &= G(x, y) := \mathrm{max}\{\langle u, Ay\rangle - \langle x, Av\rangle |
(u,v) \in Q_1 \times Q_2\}.
\end{split}
  \label{eq:dgap}
\end{eqnarray}
%}
\end{minipage}
\hspace{.5em}
\begin{minipage}{.4\linewidth}
  \includegraphics[width=1\linewidth]{grain4.png}
  \caption{A convex polytope}
  \label{fig:roi}
\end{minipage}
\end{figure}

@@

\paragraph{A note about matrix games on simplexes.} It should be
noted
that any matrix $A \in \mathbb{R}^{n_1 \times n_2}$ specifies a matrix
  game with payoff matrix $A$, for which each player's strategy
profile is a simplex; this simplex can be written in the form
\eqref{eq:polytope} by taking $E_k := (1, 1, ..., 1) \in
\mathbb{R}^{1 \times n_k}$ and $e_k = 1 \in \mathbb{R}^1$. Thus every
matrix game on simplexes can be seen as a sequential game.  Thus the
results presented in this manuscript can be trivially applied such
games in particular. Here, the polytopial $Q_k$
defined in \eqref{eq:polytope} reduce to simplexes $\Delta_{n_k}$,
and the primal-dual gap function $G(x,y)$ writes
\begin{eqnarray}G(x, y) =
\mathrm{max}\{\langle u, Ay\rangle - \langle x, Av\rangle | (u,v) \in
\Delta_{n_1} \times \Delta_{n_2}\} = \underset{0 \le i <
  n_1}{\text{max }}(Ay)_i - \underset{0 \le j < n_2}{\text{min
}}(A^Tx)_j.
\label{eq:mg_pd}
\end{eqnarray}

@@

\subsection{Our contribution}
One cannot directly attack the LCPs \eqref{eq:primal_pb} and
\eqref{eq:dual_pb} via a traditional primal-dual algorithm
(for example \cite{chambolle2010,chambolle2014ergodic}) because
computing the orthogonal projections $\proj_{Q_k}$ is very difficult
In fact, such subproblems would have to be solved
iteratively\footnote{An ``exception to the rule'' is the case where
  the $Q_k's$ are simplexes, so that the projections can be computed
  exactly using \cite{duchi2008efficient}.}. Also,
the primal-dual gap might explode even at points arbitrarily close to
the set of feasible points, leaving the algorithm with no indication
whatsoever, on whether progress is being made or not. So need to way to
\begin{itemize}
\item[(1)] avoid having to compute the projections $\proj_{Q_k}$,
\item[(2)] have control over how far we are from the set of
  equilibria and avoid infinite (and thus non-informative) primal-dual
  gaps.
\end{itemize}

Developing on an alternative notion of approximate equilibria (see
Definition \ref{thm:cool_notion})
homologous to that presented in Definition \ref{thm:approx_nash}, we
device a primal-dual algorithm that (Algorithm \ref{Tab:algo}) for
computing approximate Nash equilibria in sequential two-person
zero-sum games with incomplete information and perfect recall, and
which satisfies the above crucial requirements (1)--(2). We also
prove (Theorem \ref{thm:pd}) that --in an ergodic sense-- the number of
iterations required by the algorithm to produce an approximation
equilibrium to a precision $\epsilon$ is $\mathcal{O}(1/\epsilon)$,
with explicit values for the constants involved this worst-case cost.
These contributions will be elaborated in section \ref{sec:gsp}.
%%  The source of approximation in our scheme
%% is that we explicitly control how close we are to the feasible set
%% $Q_1 \times Q_2$.

\section{Related work}
\label{sec:related_work}
We now present a selection of algorithms that is representative of the
efforts that have been made in the literature to compute Nash
$\epsilon$-equilibria for two-person zero-sum games with incomplete
information like Texas Hold'em poker, etc.
First and foremost, let us note that for the class of games considered
here (sequential games with incomplete information), the LCPs
\eqref{eq:primal_pb} and \eqref{eq:dual_pb} are exceedingly larger
than what state-of-the-art LCP and interior-point solvers can
handle. See for example \cite{hoda2010smoothing,gilpinfirst}.

In \cite{hoda2010smoothing}, a nested iterative procedure using the
Excessive Gap Technique (EGT) \cite{nesterov2005excessive} was used
to solve the equilibrium problem \eqref{eq:gilpin}.
The authors reported a $\mathcal{O}(1/\epsilon)$ convergence rate
(which derives from the general EGT theory) for the outer-most
iteration loop.
\cite{gilpinfirst} proposed a modified verson of the technqiues in
\cite{hoda2010smoothing} and  proved a $\mathcal{O}\left(\left(\|A\| /
\delta\right) ln\left(1 / \epsilon\right)\right)$ convergence rate in
terms of the number of calls made to a first-order oracle. Here
$\delta = \delta(A, E_1, E_2, e_1, e_2) > 0$ is a certain
\textit{condition number} for the game. The crux of their technique was to
observe that \eqref{eq:gilpin} can further be written a the minimization of
the primal-dual gap function $G(x, y)$ (defined in \eqref{eq:dgap})
for the game\footnote{The minimizers of $G$ are precisely the
  equilibria of the game.}, viz
\begin{eqnarray}
\mathrm{minimize}\{G(x,y)|(x,y) \in Q_1 \times Q_2\},
\end{eqnarray}
and then show there exists a scalar
$\delta > 0$ such that for any pair of realization plans $(x, y) \in Q_1 \times Q_2$,
\begin{eqnarray}
\text{``distance between }(x, y)\text{ and the set of
equilibria'' } \le G(x, y)/\delta.
\end{eqnarray}
Their
algorithm is then derived by iteratively applying Nesterov smoothing \cite{nesterov2005a}
with a geometrically decreasing sequence of tolerance levels
$\epsilon_{n+1} = \epsilon_n / \gamma$ (with $\gamma > 1$)  $G$. It
should be noted however that
\begin{itemize}
\item[--] The constant $\delta > 0$ can be arbitrarily small, and so
  the factor $\|A\| / \delta$ in the $\mathcal{O}\left(\left(\|A\| /
\delta\right) ln\left(1 / \epsilon\right)\right)$ convergence rate can
be arbitrarily large for ill-conditioned games.
\item[--] The reported linear convergence rate is not in terms of
  basic operations (addition, multiplication, matvec, clipping, etc.),
  but in terms of the number of calls to a first-order oracle. Most
  notably, the complicated projections $\proj_{Q_k}$ are applied at
  each iteration.%%  Of
  %% course, the existence of a linear algorithm for solving
  %% \eqref{eq:gilpin} is improbable, as the usual strong-convexity
  %% conditions are absent.

@@

The primal-dual algorithm first developed in \cite{chambolle2010}, was
proposed in \cite{chambolle2014ergodic} as a way of solving matrix
games on. It should be stressed that such matrix games on simplexes are
considerably simpler than the games considered here. Indeed, the
authors in \cite{chambolle2014ergodic} used the fact that computing the
orthogonal projection of a point onto a simplex can be done in linear
time as in \cite{duchi2008efficient}. In contrast, no such efficient
algorithm is known nor is likely to exist, for the polytopes $Q_k$
defined in \eqref{eq:polytope}.%% , the strategy profiles for players in
%% the games considered here.
That notwithstanding, such projections can still be done iteratively
using for example, the algorithm in proposition 4.2 of
\cite{combettes2010dualization} or the algorithms developed in
\cite{tran2015splitting}. Unfortunately, as with any nested iterative
scheme, one would have to solve this sub-problem with finer and finer
precision.

@@

\begin{proof}
It suffices to show that at any point $(y, p, x,
q) \in S_1 \times S_2$, the primal-dual gap between the primal
LCP \eqref{eq:primal_pb} and the dual LCP \eqref{eq:dual_pb} equals
the primal-dual gap of GSP($\Psi_1$, $\Psi_2$, $g_1$, $g_2$).
Indeed, the unconstrained objective in \eqref{eq:primal_pb}, say
$a(x,y)$, can be computed as
\begin{eqnarray*}
  \begin{aligned}
    a(y,p) &= \langle e_1,p\rangle + i_{y\ge 0} + i_{-Ay + E_1^Tp \ge 0} +
    i_{E_2y = e_2}\\
    &= g_1(y,p) + \underset{x' \geq
      0}{\text{max}}\text{ }\langle x',Ay - E_1^Tp\rangle +
    \underset{q'}{\text{max}}\text{ }\langle q',E_2y - e_2\rangle\\
    &= g_1(y,p) + \underset{x',
      q'}{\text{max}}\text{ }\langle x',Ay\rangle - \langle x',
    E_1^Tp\rangle + \langle q',E_2y\rangle -
    (i_{x' \ge 0} + \langle e_2,q\rangle)\\
    &= g_1(y,p)
      - \underset{x',q'}{\text{min}}\text{ }\Psi_2(y, p, x', q') + g_2(x',
      q') = g_1(y,p)
      - \underbrace{\underset{x',q'}{\text{min}}\text{
        }\hat{\Psi}_2(y, p, x', q')}_{\phi_2(y,p)}\\
      &= g_1(y, p) - \phi_2(y, p).
  \end{aligned}
  \label{eq:a}
\end{eqnarray*}

@@

The interested reader should refer to \cite{combettes2011proximal} and
the references therewithin, for a more elaborate exposition on
proximal calculus and its use in modern convex-optimization.

\subsection{A more comfortable notion of approximate Nash equilibrium}
\begin{definition}
  Given a scalar $\epsilon > 0$ and a function $f:\mathbb{R}^n
  \rightarrow [-\infty,+\infty]$, the $\epsilon$-enlarged subgradient (or
  $\epsilon$-subgradient, for short) of $f$ is the set-valued function
  \begin{eqnarray}
\partial_\epsilon f: \mathbb{R}^n \rightarrow
2^{\mathbb{R}^n}, \hspace{.5em} x \mapsto \{s \in \mathbb{R}^n | f(z)
\ge f(x) + \langle s, z - x\rangle - \epsilon,\forall z \in
\mathbb{R}^n\}.
\end{eqnarray}
\end{definition}
The idea behind $\epsilon$-subgradients is the following. Say we wish
to minimize the function $f$. Replace the usual necessary
condition ``$0 \in \partial f(x)$'' for the optimality of $x$ with the
weaker condition ``$\partial_\epsilon f(x)$ contains a sufficiently
small vector $v$''. In fact, it is easy to see that, for each
point $x \in \mathbb{R}^n$, we have $\underset{\epsilon \rightarrow
  0^+}{\text{lim }}\partial_\epsilon f(x) = \partial f(x)$.

@@

\subsection{The proposed algorithm}
\label{sec:algo}
We now derive the algorithm (Algorithm \ref{Tab:algo}) which is the
  main object of this
manuscript, and establish the theoretical properties.

\begin{algorithm}
\caption{Primal-dual algorithm for computing approximate Nash
  Equilbria in two-person zero-sum games with incomplete information
  and perfect recall}
\label{Tab:algo}
\begin{algorithmic}[1]
\Require $\epsilon > 0$; $(y^{(0)},p^{(0)},x^{(0)},q^{(0)})
\in \mathbb{R}^{n_2}
  \times \mathbb{R}^{l_1} \times \mathbb{R}^{n_1} \times
  \mathbb{R}^{l_2}$.
\Ensure A Nash $(\epsilon,0)$-equilibrium
$({y^*},{p^*},{x^*},{q^*}) \in S_1 \times S_2$ for
the GSP \eqref{eq:unconstrained_pb}.
\State  $\lambda \leftarrow 1/\|K\|$, ${v}_a^{(0)} \leftarrow 0$, $k
\leftarrow 0$
\While{$\|{v}_{a}^{(k)}\| \ge \epsilon$}
\State $y^{(k + 1)} \leftarrow (y^{(k)} - \lambda (A^Tx^{(k)} +
E_2^Tq^{(k)}))_+$, \hspace{.5em}$p^{(k+1)} \leftarrow p^{(k)} -
\lambda(e_1-E_1x^{(k)})$
\State $x^{(k + 1)} \leftarrow (x^{(k)} + \lambda (Ay^{(k+1)} -
E_1^Tp^{(k+1)}))_+$, \hspace{.5em}$\Delta x^{(k+1)} \leftarrow
x^{(k+1)}-x^{(k)}$
\State $\Delta q^{(k+1)} \leftarrow \lambda (E_2y -
e_2)$, \hspace{.5em}$q^{(k+1)} \leftarrow q^{(k)} + \Delta q^{(k+1)}$
\State $y^{(k+1)} \leftarrow y^{(k+1)} - \lambda (A^T\Delta x^{(k+1)}
+ E_2^T\Delta q^{(k+1)})$, \hspace{.5em}$\Delta y^{(k+1)} \leftarrow
y^{(k+1)}-y^{(k)}$
\State $p^{(k+1)} \leftarrow p^{(k+1)} + \lambda E_1\Delta x^{(k+1)}$,
\hspace{.5em} $\Delta p^{(k+1)} \leftarrow p^{(k+1)}-p^{(k)}$
\State ${v}_a^{(k+1)} \leftarrow \frac{1}{\lambda}\frac{1}{k+1}(y^{(k+1)} -
y^{(0)}, p^{(k+1)} - p^{(0)}, x^{(k+1)} - x^{(0)}, q^{(k+1)} - q^{(0)})$ 
\State $k \leftarrow k + 1$
\EndWhile
\end{algorithmic}
\end{algorithm}

\begin{theorem}[Ergodic $\mathcal{O}(1/\epsilon)$ convergence]
Let $d_0$ be the orthogonal distance between the starting point
$(y^{(0)},p^{(0)},x^{(0)},q^{(0)})$ of Algorithm \ref{Tab:algo} and the
set of equilibria for the GSP \eqref{eq:unconstrained_pb}.
Then given any $\epsilon > 0$, there exists an index
$k_0 \le \frac{2d_0\|K\|}{\epsilon}$ such that after $k_0$ iterations
the algorithm produces the following a quadruplet
$(y^{k_0},p^{k_0},x^{k_0},q^{k_0})$ and a vector $v_a^{k_0}$ such that
\begin{eqnarray}
\|v_a^{k_0}\| \le \epsilon \text{ and}\hspace{.5em} v_a^{k_0} \in
\partial[\hat{\Psi}_1(., ., x^{k_0}, q^{k_0}) +
  \hat{\Psi}_2(y^{k_0}, p^{k_0}, ., .)](y^{k_0},p^{k_0},x^{k_0},q^{k_0}).
\end{eqnarray}

@@

\section{Concluding remarks}
Making use of the sequence-form representation
\cite{koller1992complexity,von1996efficient,vonequilibrium}, we have
deviced a simple and efficient primal-dual algorithm for computing
Nash equilibria in two-person zero-sum sequential games with
incomplete information (like Texas Hold'em, etc.). Our algorithm is
simple to implement, with a low constant cost per iteration, and
enjoys a rigorous convergence theory with a proven
$\mathcal{O}(1/\epsilon)$ convergence in terms of basic operations
(matvec products, clipping, etc.), to a Nash
$(\epsilon,0)$-equilibrium of the game.

Equilibrium problems are saddle-point convex-concave problems, and as
such a natural choice for algorithms for solving them would be in the
family of primal-dual algorithms. We believe such primal-dual
schemes will receive more attention in the algorithmic game theory
community in future.
